---
title       : Vacation Savings Calculator 
subtitle    : How and why to use the Vacation Savings Calculator app
author      : Matthew Fergusson
job         : Assignment for Developing Data Products class through Johns Hopkins
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Why use the Vacation Savings Calculator

#### Have you ever thought about a trip you saw in a magazine, but wondered whether or not you could ever really afford it?


#####  If you have then this is your opporunity to see what it would take to make that dream a reality.


#### All you have to do is:

1. Decide when you want to take the vacation
2. Estimate how much the vacation will cost
3. Know a little bit about your own financial and credit situation

--- .class #id 

## Step 1: Entering Your Information

#### For an example lets start off with some sample Inputs.

##### Below are sample inputs that reflect the actual entry fields of the app

```{r}
Vacation_Cost <- 3000  #Cost of Vacation
Weeks_till_Trip <- 25  #Weeks Until the Planned Trip
Cash_in_Bank <- 500    #Cash in the bank that can be used for the trip
Total_Credit <- 5000   #Total Credit from All Credit Lines
Credit_Used <- 250     #Total Credit Used from All Credit Lines
Pct_Credit_Used <- 25  #The % of your total credit that you are willing to use for vacation
Weekly_Earnings <- 500 #Weekly takehome pay after pre-tax deductions
Monthly_Rent <- 1000   #Monthly rent or mortgage payment
Other_Exp <- 500       #Other Monthly Mandatory Expenses
Pref_Budget <- 250     #How much you normally spend in a week
```

--- .class #id 

## Step 2: What is being calculated

#### The app calculates your maximum weekly budget for normal spending under 3 scenarios to pay for vacation:

1. Maxing out your credit card 
2. Using only the amount of credit you are comfortable with
3. not using any credit at all for the vacation

``` {r}
#1)
MaxOUT_Credit <- ((Total_Credit - Credit_Used) + Cash_in_Bank + Weekly_Earnings * Weeks_till_Trip - (Monthly_Rent * (12/52) + Other_Exp * (12/52)) * Weeks_till_Trip) / Weeks_till_Trip - Vacation_Cost / Weeks_till_Trip
#2)
Comfortable_Credit <- (max(Total_Credit * (Pct_Credit_Used/100) - Credit_Used,0) + Cash_in_Bank + Weekly_Earnings * Weeks_till_Trip -  (Monthly_Rent * (12/52) + Other_Exp * (12/52)) * Weeks_till_Trip)/Weeks_till_Trip - Vacation_Cost/Weeks_till_Trip
#3)
No_Credit <- (Cash_in_Bank + Weekly_Earnings * Weeks_till_Trip - (Monthly_Rent * (12/52) + Other_Exp * (12/52)) * Weeks_till_Trip)/Weeks_till_Trip - Vacation_Cost/Weeks_till_Trip
```

```{r echo=FALSE}
Result_Summary <- if(No_Credit >= Pref_Budget) {
                    "You are in the clear. You don't have to change a thing to take that vacation."
                 } else if ( Comfortable_Credit >= Pref_Budget) {
                    "You can afford the vacation while staying within the amount of credit you are willing to use; however, be careful not to rack up too much credit card debt."
      } else if (MaxOUT_Credit >= Pref_Budget) {
        "You can afford the vacation, but only by using more credit than you are comfortable with. The YOLO attitude is in vogue, but be careful of the financial hole you may be digging."        
      } else if (MaxOUT_Credit >= 0) {
        "You can barely afford the vacation. You will have to cut back your weekly budget. The YOLO attitude is in vogue, but now you are starting to get reckless."
      } else {paste("Sorry you are out of luck. You would need at least $",as.character(round(MaxOUT_Credit*Weeks_till_Trip,2)), " of additional credit while cutting your weekly spending to $0 to make this remotely possibility. If you are thinking that way your money might be better spent at a psychiotrists office. Keep dreaming, but you need to make some big changes before you can take this one on.") }
```

--- .class #id 


## Step 3: Explaining the results

#### Here you can see what the output of the app will be given the previous inputs

1. Maxing out your credit card 
```{R echo=FALSE}
MaxOUT_Credit
```
2. Using only the amount of credit you are comfortable with
```{R echo=FALSE}
Comfortable_Credit
```
3. not using any credit at all for the vacation
```{R echo=FALSE}
No_Credit
```

#### Summary:
```{R echo=FALSE}
Result_Summary
```

--- .class #id 

## Step 4: Flying, Driving, or Sailing Away to Your Dream Come True

### With the information provided by this app you will be off to visit whatever place you may desire. That is asa long as it's in the budget

### Thank you for your attention, please let me know if you have any questions